Playing with PCAs

### 2023-12-06 3-reports/playing_with_pcas.Rmd
clean up PCA, remove GWC and DIC

```{r}

data_combined_subset = make_data_subset(data_combined) %>% filter(!analysis %in% "LOI")
data_combined_wide = make_data_wide(data_combined_subset, sample_key)
data_combined_wide_NO_DIC = make_data_wide(data_combined_subset %>% 
                                             filter(!grepl("dic", name, ignore.case = TRUE)),
                                           sample_key)

fit_pca_function = function(dat){

@@

  list(num = num,
       grp = grp,
       pca_int = pca_int)
}
```

@@

```{r}
## PCA input files ----
pca_overall = fit_pca_function(data_combined_wide_NO_DIC) 
pca_overall_wle = fit_pca_function(data_combined_wide_NO_DIC %>% filter(region == "WLE"))
pca_overall_cb = fit_pca_function(data_combined_wide_NO_DIC %>% filter(region == "CB"))
```

```{r}
## PCA plots overall ----
#gg_pca_overall = 
  ggbiplot(pca_overall$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall$grp$region), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(#shape = pca_overall$grp$transect,
               color = groups))+ 
  scale_color_manual(values = c("#16879C", "#BB281E" ))+
  #scale_shape_manual(values = c(21, 19))+
  #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
  # xlim(-4,20)+
  # ylim(-8,8)+
  labs(shape="",
       title = "Overall PCA, both regions",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL
```

>>>>

```{r, eval=FALSE}

data_combined_subset = make_data_subset(data_combined) %>% filter(!analysis %in% "LOI")
data_combined_wide = make_data_wide(data_combined_subset, sample_key)
data_combined_wide_NO_DIC = make_data_wide(data_combined_subset %>% 
                                             filter(!grepl("dic", name, ignore.case = TRUE)),
                                           sample_key)
```

```{r}
## Set df for PCA

data_wide_PCA = data_combined_wide %>% dplyr::select(-c("dic", "gwc"))

```{r}
fit_pca_function = function(dat){

@@

  list(num = num,
       grp = grp,
       pca_int = pca_int)
}

@@

```{r}
## PCA input files ----
pca_overall = fit_pca_function(data_wide_PCA) 
pca_overall_wle = fit_pca_function(data_wide_PCA %>% filter(region == "WLE"))
pca_overall_cb = fit_pca_function(data_wide_PCA %>% filter(region == "CB"))
```

```{r}
## PCA plots overall ----
#gg_pca_overall = 
  ggbiplot(pca_overall$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall$grp$region), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(#shape = pca_overall$grp$transect,
               color = groups))+ 
  scale_color_manual(values = c("#16879C", "#BB281E" ))+
  #scale_shape_manual(values = c(21, 19))+
  #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
  # xlim(-4,20)+
  # ylim(-8,8)+
  labs(shape="",
       title = "Overall PCA, both regions",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL

  ggbiplot(pca_overall$pca_int, obs.scale = 1, var.scale = 1,
           groups = as.character(pca_overall$grp$transect), 
           ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = pca_overall$grp$region,
               color = pca_overall$grp$transect))+ 
 # scale_color_manual(values = c("#16879C", "#BB281E" ))+
  #scale_shape_manual(values = c(21, 19))+
  #scale_shape_manual(values = c(21, 21, 19), name = "", guide = "none")+
  # xlim(-4,20)+
  # ylim(-8,8)+
  labs(shape="",
       title = "Overall PCA, both regions",
       subtitle = "Surface horizons only")+
  theme_kp()+
  NULL
